### 2022-07-07 chapters/benchmarking/conclusion.tex
fix: bug when it incliudes a graphic
proof read :  chapter benchmarking

\section{Perspectives}
By the end of this study, we have gathered enough guidelines to make the tests more reproducible, accurate.
We created a set of new tests named \textbf{energy tests} which are more similar to performance tests.
Thanks to the work of two interns [ mamadou and adrien] we created a CI/CD platform to measure the energy consumption of Java projects and we could track the evolution of the this energy across different stages of the project.
In the figure below we see an example of this plugin.
For more details please visit the gitlab repository ... add link.

>>>>

\section{Perspectives / representativeness}
By the end of this study, we have gathered enough guidelines to make the tests more reproducible, accurate and extensible.
We created a set of new tests named \textbf{energy tests} which are more similar to performance tests.
Thanks to the work of two interns [Mamadou and Adrien] we created a CI/CD platform to measure the energy consumption of Java projects and we could track the evolution of the this energy across different stages of the project.
In the figure below we see an example of this plugin.
For more details please visit the gitlab repository\footnote{\url{https://gitlab.inria.fr/mamdiall/J-Joules}}~\footnote{\url{https://gitlab.inria.fr/mamdiall/jjoules-plugin}} .

### 2022-07-07 chapters/benchmarking/draft.tex
proof read :  chapter benchmarking

\section{Motivation}
Empirical measurements are critical tasks to capture the effect of developers' choice on software energy consumption.
To accomplish this, one should not overlook the benchmarking pitfalls highlighted in the state of the art.
Moreover, when it comes to measuring the energy consumption of software, other obstacles arise, such as the impact of hardware, the difference between the machines used for measurements purpose and the ones that may apply the findings, etc. 
Moreover, significant improvements in computer science have resulted in an increase in the quantity of obsolete results.
It is even more difficult to address all of them in studies that compare multiple solutions.
Furthermore, there may have been the appearance of new candidates, as well as the development of others between the preliminary experiment and the published results.
As a result, it is not only required to ensure the "reproducibility" of the results, but also to make way for new candidates.
Hence, we would like to propose a new concept to leverage experiments, which we call "Extension".
The capacity to provide the tools required not only to repeat the experiments, but also to include additional candidates, workloads, or relevant key performance indicators.

This chapter discusses how to solve the challanges of empirical analysis when it comes to the energy consumption experiments,

first we will dicuss how to ensure the three aspects of successful benchmark within the energy consumption. 
So section~\ref{sec:benchmarking_reproducibility} will focus on the "reproducibility" aspect and how to provide reproducible experiments without impacting the energy measurement. 
Section~\ref{sec:taming-the-energy-variation} will be about the accuracy when it comes to software energy consumption. and finally 
Section~\ref{sec:extension} will be about the tools to include new candidates, workloads, or relevant key performance indicators.

As for "representativeness" of the benchamrks, we will discuss it further in the Perspectives section.

%% this is might be the part that treats reproducibility for my case

@@

% \begin{itemize}
%     \item textbf{expermimental protocol}: Which describes, the context and  an orchestrator that handles the experiment, this orchestrator should take as an entry 
%     \item \textbf{candidates}: a set of candidates that will be used in the experiment. The candidates should be agnostic of the experiment context and they all
% \end{itemize}
\section{Reproducibility within the context of energy}\label{sec:benchmarking_reproducibility}
% Now that we have covered the general strucutre of the protocol, this section details the specific aspects of the protocol that are relevant to the energy consumption analysis.
% We will first discuss how to insure the extension of the candidates. later we will talk about the energy metrics and how to ensure the accuracy when it comes to the measurement of the energy consumption.
This section thus discusses different ways to enforce the reproducibility of experiment by comparing different ways to encapsulate the systems-under-test.

\paragraph{Virtual Machines}
First option should be using \emph{Virtual Machines} (VM), which gives researchers the freedom to choose the most appropriate tools, software and operating system that they are the most comfortable with, without paying the price to change the actual working environment, which will give them eventually more control over the dependencies and the execution environment.Therefore, Adding a new candidate is as simple as adding a new VM to the experiment.
Moreover, using a VM addresses the \emph{replication crisis} thanks to the virtual images, as even the most complex architecture can be reproduced easily by instantiating a copy of the image. 
However, this choice comes with a certain cost.
Because of the hypervisor, software will build on two kernels---the virtual machine (guest) one and the host machine one---which will provide a noticeable overhead, and will impact the performances of the system-under-test.
Therefore, we cannot use VM for experiments that are related to performance.
Another limitation of VM is the the isolation.
While this feature prevents the experimental environment from any undesirable interference from the outside world, this interaction may be required---especially when the experiment is dependent to an external

\paragraph{Containers}
Another solution would be using something that allows us to benefit from the isolation of the host OS, while easing of replication proposed by VM, and the direct interaction with the hardware that the classical method give.
Containers offer such an advantage, while ensuring the isolation and the ease of replication for application.

Figure~\ref{environement:virtualization_technique} explains the differences in architecture between the types of virtualization and container technologies:
\begin{itemize}
    \item \textsf{Type\,1} runs directly on the hardware, it is mainly used by the cloud providers where there is no host OS, but just VM, use the open-source Xen or VMware ESX hypervisors to be executed;
    \item \textsf{Type\,2} runs over the host OS, mostly used for personal computers.
    VMware server and virtualBox are popular examples of this type, and most of the researchers experimentation are run with this type. However, due to the 2 OS, the applications tend to be slower.
    \item \textsf{Containers}. Instead of its own kernel, containers use the host kernel to run their OS, which makes them lighter, faster and use the full potential of hardware.
    For this category, one can cite \emph{Docker}, \emph{Linux LXC}, or \emph{LXD}~\cite{abuabdo_virtualization_2019}.
\end{itemize}

@@

\subsection{Docker Vs. Virtual Machine}
Despite the fact that \textsf{Type\,1} is more performant than \textsf{Type\,2}, the second one is the most used in research, as most researchers tend to conduct their experiments in their own machine.
In the other hand, Docker is the most famous container technology.
In our case, we are more prone to promote Docker for two reasons:
\begin{enumerate}
    \item we need a lightweight orchestrator to limit the overhead on energy consumption of our tests, as prior work mentioned
    \cite{van2016power}
    % TODO : ask romain about the title of the paper and the reference to the paper
    [cite Morabito (2015) and van Kessel et al. (2016)],
          % cite Power efficiency of hypervisor-based virtuali- zation versus container-based virtualization. University of Amsterdam.
    \item as we are using hardware sensors to measure the energy consumption, we need to interact with the host OS.
\end{enumerate}

Special notice to \href{https://github.com/powerapi-ng/virtualwatts}{virtualwatts}. A framework that allows us to retireve the energy consumption of a virtual machine.

\subsection{Docker \& Energy}
Now that we have chosen to go with the containers technology to encapsulate our tests.
What would be the impact of this solution on the energy consumption of our tests.

Based on the studies of \cite{santos2018does}, who analysed the impact of adding the Docker layer on the energy consumption.
In their experiment, Eddie Antonio~\emph{et~ al.} run multiple benchmarks with and without Ddocker. 
They compared the resulting energy consumption and execution time.
The first step was to see the impact of Docker deamon, while there is no work, to observe the impact of the orchestrator.
Then, they conduct their experiments with the following benchmarks:
\begin{itemize}
    \item Wordpress,
    \item Reddis,
    \item PostgreSQL.
\end{itemize}

The following figures represents the energy consumption of the system-under-test while it is idle.
As one can observe in Figure~\ref{fig:docker_idle}, Docker brings an overhead of around $1,000$ Joules.

@@

In the other hand, as one can see in Figure~\ref{fig:docker_reddis}.
Docker increased the execution time of the benchmark by 50 seconds, hence increasing the energy consumption.
The authors reported that this overhead is mostly due to the Docker deamon and not the fact that the application is in a container.
Moreover, they estimated the cost of this extra energy and it was less than 0.15\$ in the worst case, which is non-significant compared to the benefits that Docker brings for isolation and reproducibility.

To summarize, Docker-based software tends to consume more energy, because mainly they take more time to be executed.
The average power consumption is higher with only \textbf{2\,Watts}, due to the execution of the Docker deamon.
This overhead can reach up to 5\% for IO-intensive application, but it is merely noticeable when it comes to CPU- or DRMA-intesive workloads.

\begin{figure}
    \includegraphics[width=.5\linewidth]{imgs/docker_vs_vm_energy_paper/reddis_time}
    \includegraphics[width=.5\linewidth]{imgs/docker_vs_vm_energy_paper/reddis_energy}
    \caption{Execution time \& energy consumption of Redis with and without Docker~\cite{santos2018does}}\label{fig:docker_reddis}
\end{figure}

\subsection{Conclusion}
As one can see, Docker impact on energy is more constat due to the introduction of the supervisor. which will be applied to all the experimenets evenly. Therefore when it comes to comparison analysis, it will mitigate its impact autimatically. 
Moreover since we have access to the host Hardware, we dont have to worry about catching the energy consumption for the SUT. 
Therefore we will be conducting all our experiments using docker as an encapsuation technology. wich will ensure the reproducibility of our experiments with a simple and easy way.

>>>>

\section{Motivation}
Empirical measurements are critical to capture the effect of developers' choices on software energy consumption.
To accomplish this, one should not overlook the benchmarking pitfalls highlighted in the state of the art.
Furthermore, when it comes to evaluating the energy consumption of software, other challenges occur, such as the impact of hardware, the difference between the machines used for measurements and those that may apply the results, and so on.
Moreover, considerable advances in computer science have resulted in an increase in the number of obsolete results.
Addressing all of them becomes significantly more difficult in studies that compare multiple treatments.

Furthermore, between the exploratory experiment and the published results, there may have been the appearance of new candidates as well as the evolution of others.
As a result, it is not only necessary to assure the "reproducibility" of the results, but it is also essential to provide room for new candidates.
As a result, we'd like to present "Extension," a new approach for leveraging experimentation.
The ability to give not only the tools needed to replicate the tests but also to integrate more candidates, workloads, or important key performance indicators.

This chapter covers ways to overcome empirical analysis challenges in energy consumption studies. First, we'll go through the three components of a successful benchmark within the energy consumption experiment. Section 1 will therefore concentrate on the "reproducibility" issue and how to deliver reproducible experiments without interfering with the energy measurement. Section 2 will discuss accuracy in software energy consumption. Finally, Section 3 will discuss tools for introducing new candidates, workloads, or other key performance indicators.

Regarding the "representativeness" of the benchmarks, we will elaborate in the section entitled Perspectives.

This chapter covers ways to overcome empirical analysis challenges in energy consumption studies. First, we'll go through the three components of a successful benchmark within the energy consumption experiment. Section~\ref{sec:benchmarking_reproducibility} will therefore concentrate on the "reproducibility" issue and how to deliver reproducible experiments without interfering with the energy measurement. Section~\ref{sec:taming-the-energy-variation} will discuss accuracy in software energy consumption. Finally, Section~\ref{sec:extension} will discuss tools for introducing new candidates, workloads, or other key performance indicators.

Regarding the "representativeness" of the benchmarks, we will elaborate in the section entitled Perspectives.
%% this is might be the part that treats reproducibility for my case

@@

% \begin{itemize}
%     \item textbf{expermimental protocol}: Which describes, the context and  an orchestrator that handles the experiment, this orchestrator should take as an entry 
%     \item \textbf{candidates}: a set of candidates that will be used in the experiment. The candidates should be agnostic of the experiment context and they all
% \end{itemize}
\section{Reproducibility within the context of energy}\label{sec:benchmarking_reproducibility}
% Now that we have covered the general strucutre of the protocol, this section details the specific aspects of the protocol that are relevant to the energy consumption analysis.
% We will first discuss how to insure the extension of the candidates. later we will talk about the energy metrics and how to ensure the accuracy when it comes to the measurement of the energy consumption.
This section discusses various techniques to encapsulate the systems-under-test in order to assure the reproducibility of an experiment.

\paragraph{Virtual Machines}
FThe very first choice should be to use \emph{Virtual Machines} (VM), which allow researchers to select the most appropriate tools, software, and operating system that they are most comfortable with without incurring the cost of changing the actual working environment, giving them more control over dependencies and the execution environment.
Furthermore, adopting a VM addresses the \emph{replication crisis} since virtual images allow even the most sophisticated architecture to be simply replicated by instantiating a copy of the image.

This option, however, comes at a cost.

Because of the hypervisor, software will be built on two kernels: one for the virtual machine (guest) and one for the host machine, resulting in a visible overhead and a negative influence on the performance of the system-under-test.

As a result, we cannot use VM for performance-related tests.

Isolation is another drawback of VM.

While this feature protects the experimental setting from unwanted interference from the outside world, it is possible that this interaction is required—-especially if the experiment is dependent on an external source such as energy montitors .

\paragraph{Containers}
Another option would be to use something that allows us to benefit from the host OS's isolation while simultaneously simplifying replication as proposed by VM and the direct interface with the hardware provided by the traditional techniques.

Containers provide such an advantage by ensuring application separation and ease of replication.
Figure~\ref{environement:virtualization_technique}depicts the architectural differences between virtualization and container technologies. There are three main types of virtualisation.
\begin{itemize}
    \item \textsf{Type\,1}: runs on the hardware directly.
          It is primarily utilized by cloud providers where there is no host OS and only VMs that run on the open-source Xen or VMware ESX hypervisors.

    \item \textsf{Type\,2}: runs on top of the host operating system and is most commonly found on personal computers.

          VMware server and virtualBox are notable examples of this category, and most researchers' experiments use them.
          However, because of the two operating systems, the applications are typically slower.

    \item \textsf{Containers}: run their operating systems on the host kernel rather than their own , which makes them smaller, faster, and more efficient in terms of hardware utilization.
          one can cite \emph{Docker}, \emph{Linux LXC}, or \emph{LXD}~\cite{abuabdo_virtualization_2019}.
\end{itemize}

@@

\subsection{Docker Vs. Virtual Machine}
Despite the fact that Type 1 is more performant than Type 2, the latter is the most used in research, as most researchers tend to conduct their experiments on their own machines. Docker, on the other hand, is the most well-known container technology. In our case, we are more likely to promote Docke for two reasons:
\begin{enumerate}
    \item As previously stated [143] [reference Morabito (2015) and van Kessel et al. (2016)], we require a lightweight orchestrator to limit the overhead on energy usage of our experiments
          \cite{van2016power}
          % TODO : ask romain about the title of the paper and the reference to the paper
          [cite Morabito (2015) and van Kessel et al. (2016)],
          % cite Power efficiency of hypervisor-based virtuali- zation versus container-based virtualization. University of Amsterdam.
    \item We need to communicate with the host OS because we are using hardware sensors to measure the energy consumption..
\end{enumerate}

A special notice to \href{https://github.com/powerapi-ng/virtualwatts}{virtualwatts}. A framework that enables us to retrieve the energy consumption of a virtual machine.

\subsection{Docker \& Energy}
Now that we've decided to use container technology to enclose our tests, what effect will this have on the amount of energy our tests use?
%%%%%%%%%

Using research from \cite{santos2018does} who examined how adding the Docker layer affected energy consumption.
\citeauthor{santos2018does}~\emph{et~ al.}conducted their experiment by running numerous benchmarks both with and without Ddocker.They contrasted the energy usage and execution time that resulted.
The first step was to observe the effects of the orchestrator and the Docker deamon while there was no work to be done.
Then, they use the following benchmarks in their experiments:
%%

\begin{itemize}
    \item Wordpress,
    \item Reddis,
    \item PostgreSQL.
\end{itemize}

The values below show the system under test's energy consumption while it is idle.
Docker has an overhead of roughly $1,000$  joules, as seen in Figure~\ref{fig:docker_idle}

Yet, as seen in Figure 3.3, Docker increased the execution duration of the benchmark by 50 seconds, increasing energy usage.
According to the authors, this overhead is primarily caused by the Docker deamon and not by the fact that the

@@

Yet, as seen in Figure~\ref{fig:docker_reddis}, Docker increased the execution duration of the benchmark by 50 seconds,which led to a significant rise in the energy usage.
According to the authors, this overhead is primarily caused by the Docker deamon and not by the fact that  the application is in a container.

Furthermore, they calculated the cost of this extra energy, which was less than 0.15\$ in the worst-case scenario, which is insignificant in comparison to the benefits of Docker for isolation and reproducibility.

To summarize, Docker-based software tendst to consume more energy since it takes longer to execute.
The execution of the Docker deamon causes an increase in average power consumption of only \textbf{2\,Watts}.
This overhead can reach up to 5\% in IO-intensive applications, while it is barely visible in CPU- or DRAM-intensive workloads.

\begin{figure}[!bht]
    \includegraphics[width=.5\linewidth]{imgs/docker_vs_vm_energy_paper/reddis_time}
    \includegraphics[width=.5\linewidth]{imgs/docker_vs_vm_energy_paper/reddis_energy}
    \caption{Execution time \& energy consumption of Redis with and without Docker~\cite{santos2018does}}\label{fig:docker_reddis}
\end{figure}

\subsection{Conclusion}
As can be seen, the introduction of the supervisor has increased the impact of Docker on energy.
which will be applied equitably to all experiments.
Therefore when it comes to comparison analysis, it will mitigate its impact autimatically.
Furthermore, because we have access to the host Hardware, we do not need to worry about capturing the SUT's energy use.
As a result, we will use Docker as an encapsuation solution for all of our experiments.
It would ensure the reproducibility of our experiments in a simple and straightforward manner.
\clearpage

### 2022-07-07 chapters/benchmarking/experiment.tex
proof read :  chapter benchmarking

% To satisfy those three requirements. we elaborated a benchmarking protocol that aims to help reaserchers and practicioners in their experementations.

% for each aspect we present a set of guidelines and tools.
%% This is a conclusion maaan  !! 
\section{Taming the energy variation for the benchmarking protocol}\label{sec:taming-the-energy-variation}

\subsection{Objective}
While the previous part aims to create a large umbrella that covers all empirical benchmarks related to computer science, we rather aim to narrow our study to energy benchmarking only.
In other words, we will focus on pitfalls that hinders the energy claims of software.
We are clearly aware of the impact of hardware on energy variations.
However, we believe that there is still a room to control this energy variation for practitioners by using only parameters that can be tuned.
To do so, we have inducted a set of empirical experiments using the guidelines provided by state of the art, to determine which controllable factors can reduce the variation of the energy consumption of benchmarks.
% We will first start with evaluating the parameters mentioned by \cite{heinrich_predicting}.

In this part we will try to answer the following Research questions.

\begin{description}
    \item[\textsc{RQ}~1:] Does the benchmarking protocol affect the energy variation?
    \item[\textsc{RQ}~2:] How important is the impact of the processor features on the energy variation?
    \item[\textsc{RQ}~3:] What is the impact of the operating system on the energy variation? and finally
    \item[\textsc{RQ}~4:] Does the choice of the processor matter to mitigate the energy variation?
\end{description}

\subsection{Experimental Setup}\label{subsec:setup}
This section describes our detailed experimental environment, covering the clusters configuration and the benchmarks we used  justifying our experimental methodology.

\subsubsection{Measurement Context}
There are three main contexts.
\begin{itemize}
    \item Different machines with different configuration
    \item Different machines with the same configuration
    \item Same machine
\end{itemize}

To satisfy those requirements we have used the platform Grid5000 (G5K)~\cite{grid5000,margery:hal-00965708},a large-scale and flexible testbed for experiment-driven research distributed across all france.
Grid5000 offers  multiple clusters composed with 4 up to 124 identical machine with different configurations for each cluster.
For our experiment, we have considered 4 clusters.
Our main criterions was the CPU Configuration.
the table below~\ref{table:g5k}, presents a description of the 4 clusters.

@@

As most of the nodes are equipped with two sockets (physical processors), we use the acronym \textsf{CPU} or \textsf{socket} to designate one of the two sockets and \textsf{PU} for the \emph{processing unit}. For our study we consider hyper-threads as distinct \textbf{PU}

As an example, Figure~\ref{fig:topo} illustrates a detailed topology of a node belonging to the cluster \textsf{Dahu}.

@@

First, \textbf{scalability} : We wanted to gather the highest possible amount insights (regarding time spent for the experiment, therefore we wanted some benchmarks who can scale according to the number of PUs and can fulfill different scenarios.
Second criterion is \textbf{representativeness} : as mentioned in the challenges, a workload has to be representative otherwise the experiment would be inconsistent \cite{stephen_evaluate_2012}.
To satisfy those criterions we went through state of the art and looked for the most used benchmarks for testing the performance of the hardware, and then we selected the scalable ones.
Our candidate is  emph{NAS Parallel Benchmark} (NPB v3.3.1)~\cite{Bailey:1991:NPB:125826.125925}: one of the most used benchmarks for \emph{HPC}. We used the applications (\textsf{LU}), the \emph{Conjugate Gradient} (\textsf{CG}) and \emph{Embarrassingly Parallel} (\textsf{EP}) computation-intensive benchmarks in our experiments, with the \textsf{C} data class.
Further more we have used other applications to validate our results using more applications such as \texttt{Stress-ng v0.10.0},\footnote{\url{https://kernel.ubuntu.com/~cking/stress-ng}} \texttt{pbzip2 v1.1.9},\footnote{\url{https://launchpad.net/pbzip2/}} \texttt{linpack}\footnote{\url{http://www.netlib.org/linpack}} and \texttt{sha256 v8.26}\footnote{\url{https://linux.die.net/man/1/sha256sum}}.

\subsubsection{Metrics \& Measurement Tools}
Our metric for the accuracy of the test is the Standard deviation aka \textbf{STD} of the energy consumption. Therefore wether the tests consumes more or less energy is out of our scope.
To study this variation we need first a tool to measure the energy consumption. For this we used \textsc{PowerAPI}~\cite{DBLP:journals/jss/ColmantRKSFS18}, which is a power monitoring toolkit that is based on \emph{Intel Running Average Power Limit} (RAPL) \cite{Khan:2018:RAE:3199681.3177754}. The advantage of PowerAPI is that it reports the Energy consumption of CPU and DRAM at a socket level.

@@

\begin{table}[h!]
    \centering
    \caption{Experimental Guidelines for Energy Variations}
    \small
    \begin{tabular}{|p{4.7cm}|c|c|}
        \hline
        \textbf{Guideline}                                                                                                                                                  & \textbf{Load} & \textbf{Gain}     \\
        \hline
        \hline
        Use a low TDP CPU        & Low \& medium & Up to $3\times$   \\
        \hline
        Disable the CPU C-states & Low           & Up to $6\times$   \\
        \hline
        Use the least of sockets in a case of multiple CPU   & Medium        & Up to $30\times$\\
        \hline
        Avoid the usage of hyper-threading whenever possible & Medium        & Up to $5\times$\\
        \hline
        Avoid rebooting the machine between tests  & High & Up to $1.5\times$ \\
        \hline
        Do not relate to the machine idle variation to isolate a test EC, the CPU/OS changes its behavior when a test is running and can exhibit less variation than idle & Any & ---\\
        \hline
        Rather focus the optimization efforts on the system under test than the OS & Any           & ---               \\
        \hline
        Execute all the similar and comparable experiments on a same machine. Identical machines can exhibit many differences regarding their energy behavior             & Any           & Up to $1.3\times$ \\
        \hline
    \end{tabular}
    \label{table:guidelines}
\end{table}

>>>>

\newpage
% To satisfy those three requirements. we elaborated a benchmarking protocol that aims to help reaserchers and practicioners in their experementations.
% for each aspect we present a set of guidelines and tools.
%% This is a conclusion maaan  !! 
\section{Improving accuracy by Taming the energy variation}\label{sec:taming-the-energy-variation}

While the previous section aimed to ensure the reproducibility of our software energy consumption experiments, this section will provide a collection of tips and tools to help you increase the accuracy of these experiments.
We are well aware of the effect that hardware has on energy fluctuations.
However, we feel that there is still an opportunity for practitioners to minimize this energy variance by employing solely tuneable factors.
To that end, we conducted a series of empirical studies utilizing state-of-the-art recommendations to discover which controllable elements can limit the variations of benchmark energy usage.
% We will first start with evaluating the parameters mentioned by \cite{heinrich_predicting}.

In this section we will try to answer the following Research questions.
\begin{description}
    \item[\textsc{RQ}~1:] Does the benchmarking protocol affect the energy variation?
    \item[\textsc{RQ}~2:] How important is the impact of the processor features on the energy variation?
    \item[\textsc{RQ}~3:] What effect does the operating system have on energy variation?
    \item[\textsc{RQ}~4:] Does the choice of processor make a difference in reducing the energy variation?
\end{description}

\subsection{Experimental Setup}\label{subsec:setup}
This section describes our detailed experimental environment, covering the cluster configuration and the benchmarks we used to justify our experimental methodology.

\subsubsection{Measurement Context}
There are three main contexts.
\begin{itemize}
    \item different machines with different settings;
    \item different machines with the same settings ;
    \item the same machine. 
\end{itemize}

We used the platform Grid5000 (G5K)~\cite{grid5000,margery:hal-00965708}, which is a large, flexible testbed for experiment-driven research that is spread out across all of France, to meet these needs.Grid5000 has several clusters that are made up of 4 to 124 identical machines with different configurations for each cluster.For our experiment, we looked at four groups. Our main criterion was the CPU configuration.
The table below~\ref{table:g5k}, presents a description of the four cluster that have been chosen for our experiments.

@@

As most of the nodes are equipped with two sockets (physical processors), we use the acronym \textsf{CPU} or \textsf{socket} to designate one of the two sockets and \textsf{PU} for the \emph{processing unit}. For our study we consider hyper-threads as distinct \textbf{PUs}.

Figure~\ref{fig:topo}  shows the detailed topology of a node in the cluster \textsf{Dahu}. as an example.

@@

First, \textbf{scalability} : We wanted to gather the highest possible amount insights (regarding time spent for the experiment, therefore we wanted some benchmarks who can scale according to the number of PUs and can fulfill different scenarios.
Second criterion is \textbf{representativeness} : as mentioned in the challenges, a workload has to be representative otherwise the experiment would be inconsistent \cite{stephen_evaluate_2012}.
To satisfy those criterions we went through state of the art and looked for the most used benchmarks for testing the performance of the hardware, and then we selected the scalable ones.
%%%%%% 
We picked the benchmarks based on two criteria.

First, \textbf{scalability}: We wanted to learn as much as possible about the experiment in the time we had, so we needed some benchmarks that could grow or shrink with the number of PUs and work in different situations.

The second criterion is whether or not the workload is \textbf{representative}. As stated in the challenges, a workload must be representative or the experiment would be inconsistent~\cite{stephen_evaluate_2012}.

To meet these criteria, we looked at the "state of the art" and found the most common benchmarks used to test the performance of hardware. We then chose the ones that could be scaled up. 
Our candidate is  emph{NAS Parallel Benchmark} (NPB v3.3.1)~\cite{Bailey:1991:NPB:125826.125925}: one of the most used benchmarks for \emph{HPC}. We used the applications (\textsf{LU}), the \emph{Conjugate Gradient} (\textsf{CG}) and \emph{Embarrassingly Parallel} (\textsf{EP}) computation-intensive benchmarks in our experiments, with the \textsf{C} data class.
Further more we have used other applications to validate our results using more applications such as \texttt{Stress-ng v0.10.0},\footnote{\url{https://kernel.ubuntu.com/~cking/stress-ng}} \texttt{pbzip2 v1.1.9},\footnote{\url{https://launchpad.net/pbzip2/}} \texttt{linpack}\footnote{\url{http://www.netlib.org/linpack}} and \texttt{sha256 v8.26}\footnote{\url{https://linux.die.net/man/1/sha256sum}}.

\subsubsection{Metrics \& Measurement Tools}
Our metric for the accuracy of the test is the Standard deviation aka \textbf{STD} of the energy consumption. Therefore whether the tests consumes more or less energy is out of our scope.
To study this variation we need first a tool to measure the energy consumption. For this we used \textsc{PowerAPI}~\cite{DBLP:journals/jss/ColmantRKSFS18}, which is a power monitoring toolkit that is based on \emph{Intel Running Average Power Limit} (RAPL) \cite{Khan:2018:RAE:3199681.3177754}. The advantage of PowerAPI is that it reports the Energy consumption of CPU and DRAM at a socket level.

@@

\begin{table}[h!]
    \centering
    \caption{Experimental Guidelines for Energy Variations}
    \small
    \begin{tabular}{|p{4.7cm}|c|c|}
        \hline
        \textbf{Guideline}                                                                                                                                                & \textbf{Load} & \textbf{Gain}     \\
        \hline
        \hline
        Use a low TDP CPU                                                                                                                                                 & Low \& medium & Up to $3\times$   \\
        \hline
        Disable the CPU C-states                                                                                                                                          & Low           & Up to $6\times$   \\
        \hline
        Use the least of sockets in a case of multiple CPU                                                                                                                & Medium        & Up to $30\times$  \\
        \hline
        Avoid the usage of hyper-threading whenever possible                                                                                                              & Medium        & Up to $5\times$   \\
        \hline
        Avoid rebooting the machine between tests                                                                                                                         & High          & Up to $1.5\times$ \\
        \hline
        Do not relate to the machine idle variation to isolate a test EC, the CPU/OS changes its behavior when a test is running and can exhibit less variation than idle & Any           & ---               \\
        \hline
        Rather focus the optimization efforts on the system under test than the OS                                                                                        & Any           & ---               \\
        \hline
        Execute all the similar and comparable experiments on a same machine. Identical machines can exhibit many differences regarding their energy behavior             & Any           & Up to $1.3\times$ \\
        \hline
    \end{tabular}
    \label{table:guidelines}
\end{table}
